Temporal stability in human interaction networks

### 2015-06-30 paper.tex
adicionada seção sobre o uso que fizemos do pca

\subsection{Topological metrics for Principal Component Analysis}\label{measures}

The topology of the networks was studied using Principal Component Analysis (PCA)~\cite{pca} with a small selection of the most basic and fundamental measurements for each vertex, as follows:

>>>>

\subsection{Principal Component Analysis of topological metrics}\label{sec:pca}
Principal Component Analysis (PCA) is a well documented technique~\cite{pca} and was used to to know: 1) what metrics contribute to each principal component and in which proportion; 2) how much of the dispersion is concentrated in each component; 3) expected values and dispersions for these quantities over various networks.

Be $\mathbf{X}=\{X[i,j]\}$ a matrix of all vertices $i$ and the respective
values for each metric $j$, 
$\mu_X [j]=\frac{\sum_i X[i,j]}{I}$ the mean of metric $j$ over all $I$ vertices, 
$\sigma_X [j]=\sqrt{\frac{\sum_i (X[i,j]-\mu_X [j])^2}{I}}$ the standard deviation of metric $j$,
and $\mathbf{X'}=X'[i,j]={\frac{X[i,j]-\mu_X[j]}{\sigma_X[j]}}$ 
the matrix with the \emph{z-score} of each metric. 
Be $\mathbf{V}=\{V[j,k]\}$ the matrix $J\times J$ of eigenvectors
of the covariance matrix $\mathbf{C}$
of $\mathbf{X'}$, one eigenvector per column.
Each eigenvector combines the original measures into one principal component, therefore
$V'[j,k]=100*\frac{|V[j,k]|}{\sum_{j'} |V[j',k]|}$
delivers the percentage of the principal component $k$
that is due to the measure $j$.
With $k$ eigenvectors 
$D[k]$,
it is enough to observe $D'[k]=100*\frac{D[k]}{\sum_{k'}D[k']}$
to know the percentage of dispersion that each principal component
is responsible for.
With the eigenvalues $k$ in ascending order
we observe, in general, the three first eigenvalues and
the respective eigenvectors in percentages:
$\{(V'[j,k],\;D'[k])\}$.
These usually reveal sufficient patterns for a sound analysis and sum between 60 and 95\% of the dispersion.
Namely, given $L$ observations $l$, each with $k$ pairs of
eigenvalues and eigenvectors, we observe
the mean
$\mu_{V'}[j,k]$
and the standard deviation $\sigma_{V'}[j,k]$ 
of the metric $j$ in the principal component $k$,
and the mean
$\mu_{D'}[k]$
and the standard deviation 
$\sigma_{D'}[k]$
of the contribution of the component $k$ to the dispersion
of the system:

\begin{align}\label{eq:pca}
\mu_{V'}[j,k]   &=\frac{\sum_{l=1}^L V'[j,k,l]}{L}\nonumber\\
\sigma_{V'}[j,k]&=\sqrt{\frac{\sum_{l=1}^L (\mu_{V'}-V'[j,k,l])^2}{L}}\\\nonumber
\mu_{D'}[k]&=\frac{\sum_{l=1}^L D'[k,l]}{L}\\\nonumber
\sigma_{D'}[k]&=\sqrt{\frac{\sum_{l=1}^L (\mu_{D'}-D'[k,l])^2}{L}}
\end{align}

The covariance matrix 
$\mathbf{C}$, is the correlation matrix because $\mathbf{X'}$ is normalized. Therefore, it is also directly observed as a first clue for patterns.
This was accomplished with the most simple associations:
low absolute values indicate low correlation (and a possible independence);
high values indicate positive correlation; negative values with high absolute value indicate negative correlation.

\subsubsection{Metrics considered}\label{measures}

The topology of the networks was studied using PCA with a small selection of the most basic and fundamental measurements for each vertex, as follows:
